---
title: "NC Redistricing - February Update"
description: |
  Here we go again. .
author:
  - name: Jacob Ford
    url: https://jford-dchcmpo.github.io/Blog/posts/2021-12-30-Redistricting-and-Racial-Targetting/
date: 02-20-2022
output:
  distill::distill_article:
    self_contained: false
---


Late last week, Republicans in the North Carolina General Assembly cobbled together the [political boundaries](https://thehill.com/homenews/state-watch/594908-north-carolina-legislature-approves-new-us-house-maps) that, if approved and adopted, will define the political delineations in one of the fastest growing states in the country for the next decade. In my inaugural [post](https://jford-dchcmpo.github.io/Blog/posts/2021-12-07-Redistricting-and-Racial-Targetting/), I trained a few ML algorithms to determine if the recently thrown out political lines would perform better at sorting demographic groups into either Democratic or Republican districts. I'm not breaking any barrier of knowledge in saying minorities vote more often for Democrats, though in 2020 we saw how the hispanic vote (itself a constellation of dozens of nationalities and political priorities, impossible to untangle) once though a sure win for Democrats shifted to Republicans in key races. However, one of the arguments against political gerrymandering is that with the aid of immense voter level data, including demograhpics and previous vote history, that political lines are sorting us too well. That politicians are picking the voters, not the other way around. 

With that being said, let's add in the new lines. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(caliperR)
library(nngeo)
library(tcadr)
library(sf)
library(leaflet)
library(PL94171)
library(janitor)
library(dplyr)
library(tidycensus)
library(tidyverse)
options(tigris_use_cache = TRUE)
library(sf)
library(data.table)
library(dplyr)
library(plotly)
library(randomForest)

# remotes::install_github("walkerke/tidycensus")
# remotes::install_github("walkerke/tigris")
library(tidycensus)
library(tidyverse)
library(tigris)

vars_2020 <- load_variables(2020,"pl",cache=TRUE)
vars_2010 <- load_variables(2010,"sf1", cache=TRUE)
vars_2019 <- load_variables(2019,"acs5", cache=TRUE)

trm_taz_g2 <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/Archive/TRM_G2_TAZ_CaliperUpdated.shp")



nc_congressional_2019 <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/2019 NC Congressional/NC_Congress_2019.shp") 

nc_congressional_2021 <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/2021 NC Congressional/SL 2021-174 Congress.shp")

nc_congressional_2021_feb <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/February Update/SL 2022-3.shp")

house_old<- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/NC_House_2019.shp")
house_new<- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/NC_House_2021.shp")
house_feb <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/February Update/SL 2022-4.shp")

senate_old<- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/NC_Senate_2019.shp")
senate_new<- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/NC_Senate_2021.shp")
senate_feb <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/Voting Districts/February Update/SL 2022-2.shp")

nc_congressional_2021$DISTRICT <- as.integer(nc_congressional_2021$DISTRICT)


##Change CRS


nc_congressional_2019 <- st_transform(nc_congressional_2019, st_crs(trm_taz_g2)) %>% st_make_valid()
nc_congressional_2021 <- st_transform(nc_congressional_2021, st_crs(trm_taz_g2)) %>% st_make_valid()
nc_congressional_2021_feb <- st_transform(nc_congressional_2021_feb, st_crs(trm_taz_g2)) %>% st_make_valid()

house_old <- st_transform(house_old, st_crs(trm_taz_g2)) %>% st_make_valid()
house_new <- st_transform(house_new, st_crs(trm_taz_g2)) %>% st_make_valid()
house_feb <- st_transform(house_new, st_crs(trm_taz_g2)) %>% st_make_valid()

senate_old <- st_transform(senate_old, st_crs(trm_taz_g2)) %>% st_make_valid()
senate_new <- st_transform(senate_new, st_crs(trm_taz_g2)) %>% st_make_valid()
senate_feb <- st_transform(senate_new, st_crs(trm_taz_g2)) %>% st_make_valid()




nc_congressional_2021 <- nc_congressional_2021[order(nc_congressional_2021$DISTRICT),]

nc_congressional_2021$Partisan_Lean <- c(-18,1,-20,-10,25,42,-20,-20,45,-26,-16,-16,-25,-12)

##source for paritisan lean: https://projects.fivethirtyeight.com/redistricting-2022-maps/north-carolina/
nc_congressional_2021$color <- case_when(
       nc_congressional_2021$Partisan_Lean  < -19 ~ -3,
        nc_congressional_2021$Partisan_Lean < -10 ~ -2,
        nc_congressional_2021$Partisan_Lean < -5 ~ -1,
        nc_congressional_2021$Partisan_Lean < 5 ~ 0, 
        nc_congressional_2021$Partisan_Lean  < 10 ~ 1,
        nc_congressional_2021$Partisan_Lean > 10 ~ 2 
         
)

nc_congressional_2021$Pred_Party <- case_when(
       nc_congressional_2021$Partisan_Lean  < 2 ~ 0,
        nc_congressional_2021$Partisan_Lean > 2 ~ 1
         
)

nc_congressional_2021_feb <- nc_congressional_2021_feb[order(nc_congressional_2021_feb$DISTRICT),]

nc_congressional_2021_feb$Partisan_Lean <- c(-18,1,-20,-10,25,42,-20,-20,45,-26,-16,-16,-25,-12)

##source for paritisan lean: https://projects.fivethirtyeight.com/redistricting-2022-maps/north-carolina/
nc_congressional_2021_feb$color <- case_when(
       nc_congressional_2021_feb$Partisan_Lean  < -19 ~ -3,
        nc_congressional_2021_feb$Partisan_Lean < -10 ~ -2,
        nc_congressional_2021_feb$Partisan_Lean < -5 ~ -1,
        nc_congressional_2021_feb$Partisan_Lean < 5 ~ 0, 
        nc_congressional_2021_feb$Partisan_Lean  < 10 ~ 1,
        nc_congressional_2021_feb$Partisan_Lean > 10 ~ 2 
         
)

nc_congressional_2021_feb$Pred_Party <- case_when(
       nc_congressional_2021_feb$Partisan_Lean  < 2 ~ 0,
        nc_congressional_2021_feb$Partisan_Lean > 2 ~ 1
         
)




## You can use st_join and tag each census block centroid with the party ID from congressional layer,
## but it is more efficient to tag in a separate GIS software, then read back in already tagged. Work smarter not harder. 

centroids <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/centroids.shp") %>%
  mutate(PARTY_2019 = P_2019) %>%
  filter(!(is.na(PARTY_2019)))

```


## Visualize new maps

The state house and state senate maps are final, only the congressional maps are being altered. I didn't know that until I added the new maps (February in click option) to the below. If you don't believe me, just find out for yourself. 

Recall, to train the models on the new congressional lines, I assumed the 2022 midterms would follow FiveThirtyEight's [Partisan Lean](https://projects.fivethirtyeight.com/redistricting-2022-maps/north-carolina/). This assumes the worst for Democrats in North Carolina, which I imagine is more likely than not. 

```{r echo=FALSE, message=FALSE, warning=FALSE}




congress_2019 <- st_transform(nc_congressional_2019, st_crs(trm_taz_g2)) %>% st_make_valid()

pal <- colorFactor(
  palette=c('blue','red'),
  domain=congress_2019$sParty
)



congress_2021 <- st_transform(nc_congressional_2021, st_crs(trm_taz_g2)) %>% st_make_valid()

congress_2021_Feb <- st_transform(nc_congressional_2021_feb, st_crs(trm_taz_g2)) %>% st_make_valid()

library(RColorBrewer)

colfunc <- colorRampPalette(c("#2C77BF", "red"))




pal_2021 <- colorFactor(
  palette=c('darkred','red', 'pink', 'grey', 'blue', 'darkblue'),
  domain=c(-3,-2,-1,0,1,2)
)



map <- leaflet(congress_2019) %>%
  addProviderTiles("CartoDB.Positron") %>%
  ##Congress
  addPolygons(data=congress_2019, group="Congress Old",
              popup=paste("Congressional District: ", congress_2019$District, "<br>",
                          "Party: ", congress_2019$sParty, "<br>",
                          "Representative 2019: ", congress_2019$sName),
              weight=1,
              color =~pal(congress_2019$sParty)) %>%
  addPolygons(data=congress_2021, group="Congress New",
              popup=paste("Congressional District: ", congress_2021$DISTRICT, "<br>", "Partisan Lean: ", congress_2021$Partisan_Lean ),dashArray="3", color =~pal_2021(congress_2021$color),
              weight=1, fillOpacity =0.5) %>%
  
    addPolygons(data=congress_2021_Feb, group="Congress February",
              popup=paste("Congressional District: ", congress_2021_Feb$DISTRICT, "<br>", "Partisan Lean: ", congress_2021_Feb$Partisan_Lean ),dashArray="3", color =~pal_2021(congress_2021_Feb$color),
              weight=1, fillOpacity =0.5) %>%
  
  
  ##Senate
  addPolygons(data=senate_old, group="Senate Old",
              popup=paste("Senate District: ", senate_old$District, "<br>",
                          "Party: ", senate_old$sParty, "<br>",
                          "Senator 2019: ", senate_old$sFirstName, " ", senate_old$sLastName),
              weight=1,
              color =~pal(senate_old$sParty)) %>%
  addPolygons(data=senate_new, group="Senate New",
              popup=paste("Congressional District: ", senate_new$District, "<br>" ),dashArray="3", color ='black',
              weight=2, fillOpacity =0) %>%
    addPolygons(data=senate_feb, group="Senate February",
              popup=paste("Congressional District: ", senate_feb$District, "<br>" ),dashArray="3", color ='black',
              weight=2, fillOpacity =0) %>%
  
  
  ##House
  addPolygons(data=house_old, group="House Old",
              popup=paste("House District: ", house_old$District, "<br>",
                          "Party: ", house_old$sParty, "<br>",
                          "Representative 2019: ",  house_old$sFirstName, " ", house_old$sLastName),
              weight=1,
              color =~pal(house_old$sParty)) %>%
  addPolygons(data=house_new, group="House New",
              popup=paste("Congressional District: ", house_new$District, "<br>" ),dashArray="3", color ='black',
              weight=2, fillOpacity =0) %>%
  
    addPolygons(data=house_feb, group="House February",
              popup=paste("Congressional District: ", house_feb$District, "<br>" ),dashArray="3", color ='black',
              weight=2, fillOpacity =0) %>%
  
  addLayersControl(overlayGroups = c("Congress Old", "Congress New", "Congress February", "Senate Old", "Senate New", "Senate February", "House Old", "House New", "House February"))

map %>% hideGroup("Congress New")%>% hideGroup("Senate New")%>% hideGroup("House New")%>% hideGroup("Senate Old") %>% hideGroup("House Old") %>% hideGroup("Senate February") %>% hideGroup("House February")%>% hideGroup("Congress February")



  
 


```













Told you so. 

## Summary Stats

We'll make the same summary stats to start. Notice how the February updated lines seem to be a large improvement in racial parity - Blacks were nearly twice as likely to be represented by a Democrat with the 2010-2019 congressional lines, in the February update it's almost even. This process requires tagging each of the nearly quarter million census blocks in North Carolina to the congressional district represented.


```{r echo=FALSE, message=FALSE, warning=FALSE}



centroids$Med_income_log <- case_when(
  centroids$MD_INCM == 0 ~ 0,
  centroids$MD_INCM > 0 ~ log(centroids$MD_INCM))

## 2019 
temp <- centroids %>%
  st_drop_geometry() %>%
  mutate(`PARTY 2019` = case_when(
    PARTY_2019 == 0 ~ "Republican",
    PARTY_2019 == 1 ~ "Democrat"
  )) %>%
  group_by(`PARTY 2019`) %>%
  summarize(Mean_white = mean(PCT_WHT, na.rm=T),
            Mean_black = mean(PCT_BLC, na.rm=T),
            Mean_hispanic = mean(PCT_HSP, na.rm=T),
            Mean_asian = mean(PCT_ASN, na.rm=T),
            Mean_inc = mean(MD_INCM, na.rm=T))

library(DT)

datatable(temp, caption = "Average Racial Breakdown by 2019 Congressional Representation") %>%
  formatRound('Mean_white', 3) %>%
  formatRound('Mean_black', 3) %>%
  formatRound('Mean_hispanic', 3) %>%
  formatRound('Mean_asian', 3) %>%
  formatRound('Mean_inc',2)

## 2020

temp <- centroids %>%
  st_drop_geometry() %>%
  mutate(`PARTY 2021` = case_when(
    PARTY_2021 == 0 ~ "Republican",
    PARTY_2021 == 1 ~ "Democrat"
  )) %>%
  group_by(`PARTY 2021`) %>%
  summarize(Mean_white = mean(PCT_WHT, na.rm=T),
            Mean_black = mean(PCT_BLC, na.rm=T),
            Mean_hispanic = mean(PCT_HSP, na.rm=T),
            Mean_asian = mean(PCT_ASN, na.rm=T),
            Mean_inc = mean(MD_INCM, na.rm=T))

library(DT)

datatable(temp, caption = "Average Racial Breakdown by 2021 Congressional Representation") %>%
  formatRound('Mean_white', 3) %>%
  formatRound('Mean_black', 3) %>%
  formatRound('Mean_hispanic', 3) %>%
  formatRound('Mean_asian', 3) %>%
  formatRound('Mean_inc',2)



## February

temp <- centroids %>%
  st_drop_geometry() %>%
  mutate(`February_Party` = case_when(
    PARTY_2022 == 0 ~ "Republican",
    PARTY_2022 == 1 ~ "Democrat"
  )) %>%
  group_by(`February_Party`) %>%
  summarize(Mean_white = mean(PCT_WHT, na.rm=T),
            Mean_black = mean(PCT_BLC, na.rm=T),
            Mean_hispanic = mean(PCT_HSP, na.rm=T),
            Mean_asian = mean(PCT_ASN, na.rm=T),
            Mean_inc = mean(MD_INCM, na.rm=T))

library(DT)

datatable(temp, caption = "February Update") %>%
  formatRound('Mean_white', 3) %>%
  formatRound('Mean_black', 3) %>%
  formatRound('Mean_hispanic', 3) %>%
  formatRound('Mean_asian', 3) %>%
  formatRound('Mean_inc',2)

```












```{r message=FALSE, warning=FALSE, include=FALSE}

##Train and Test Data Sets

centroids <- centroids %>% st_drop_geometry()
train <- centroids[1:177478,]
test <- centroids[177479:236638,]

```




```{r message=FALSE, warning=FALSE, include=FALSE}

### 2019 Model Outputs

model_2019_blocks <- glm(PARTY_2019 ~PCT_HSP + PCT_WHT + PCT_BLC + PCT_ASN + Med_income_log ,family=binomial(link='logit'),data=train)

summary(model_2019_blocks)





```


```{r message=FALSE, warning=FALSE, include=FALSE}
### 2021 Model Outputs


model_2021 <- glm(PARTY_2021 ~PCT_HSP + PCT_WHT + PCT_BLC + PCT_ASN + Med_income_log ,family=binomial(link='logit'),data=train)

summary(model_2021)



```

```{r message=FALSE, warning=FALSE, include=FALSE}
### February Model


feb_model <- glm(PARTY_2022 ~PCT_HSP + PCT_WHT + PCT_BLC + PCT_ASN + Med_income_log ,family=binomial(link='logit'),data=train)

summary(feb_model)



```





## Logistic Regression Curve

Lets see if the new lines hold up with the logistic model, determining how likely a particular demographic is to be represented by a Democrat (note 1 - the y-axis = probability Republican). Black and Asian populations are more likely to be represented by Republicans in large numbers in the February update. Note how at the very end of the logistic curve, in those census blocks with 90, 95% racial homogeneity, the likelihood of being represented by a democrat is far less likely compared to the 2010-2019 or 2021 proposed districts. This only holds for black populations as you can see in the below: 

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(test, aes(x=PCT_ASN, y=PARTY_2019)) + geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Asian") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts Asian")

ggplot(test, aes(x=PCT_ASN, y=PARTY_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Asian") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts Asian")

ggplot(test, aes(x=PCT_ASN, y=PARTY_2022)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Asian") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")

#=====================================================================#


ggplot(test, aes(x=PCT_BLC, y=PARTY_2019)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Black") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts Black")

ggplot(test, aes(x=PCT_BLC, y=PARTY_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Black") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts Black")


ggplot(test, aes(x=PCT_BLC, y=PARTY_2022)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Black") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")

#=====================================================================#

ggplot(test, aes(x=PCT_WHT, y=PARTY_2019)) + geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent White") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts White")


ggplot(test, aes(x=PCT_WHT, y=PARTY_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent White") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts White")

ggplot(test, aes(x=PCT_WHT, y=PARTY_2022)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent White") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")
#=====================================================================#

ggplot(test, aes(x=PCT_HSP, y=PARTY_2019)) + geom_point(fill='red') + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Hispanic") +
  scale_y_continuous(name="Percent Democrat 2019")+ggtitle("2019 Districts Hispanic")

ggplot(test, aes(x=PCT_HSP, y=PARTY_2021)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Hispanic") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("2021 Districts Hispanic")

ggplot(test, aes(x=PCT_HSP, y=PARTY_2022)) + geom_point() + 
  stat_smooth(method="glm", method.args=list(family=binomial(link='logit')), se=FALSE)+theme_minimal()+ scale_x_continuous(name="Percent Hispanic") +
  scale_y_continuous(name="Percent Democrat 2021")+ggtitle("February Update")

#=====================================================================#




```


## Model Fits

The logistic model fit in the 2010-19 districts was a paltry 71.1%, drastically improved upon by the 2021 struck-down lines - up to 88.3%. The February lines are in the middle, at 82.6%; meaning this simple logistic model, fed with nearly a quarter million census block values, developed into a machine learning algorithm that could predict with 82.6% accuracy the political representation of a district based solely on racial makeup. 



```{r echo=FALSE, message=FALSE, warning=FALSE}

## Predictive Fits

pred <- predict(model_2019_blocks, newdata = test, type = "response")



y_pred_num <- ifelse(pred > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
y_act <- test$PARTY_2019

pred_2019 <- mean(y_pred == y_act, na.rm=T) 

#=====================================================================#

pred <- predict(model_2021, newdata = test, type = "response")

y_pred_num <- ifelse(pred > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
y_act <- test$PARTY_2021

pred_2021 <- mean(y_pred == y_act, na.rm=T) 





#=====================================================================#
pred <- predict(feb_model, newdata = test, type = "response")

y_pred_num <- ifelse(pred > 0.5, 1, 0)
y_pred <- factor(y_pred_num, levels=c(0, 1))
y_act <- test$PARTY_2022

feb_pred <- mean(y_pred == y_act, na.rm=T) 
#=====================================================================#



vals <- c(pred_2019, pred_2021,feb_pred)





df <- data.frame(Year = c('2019', '2021', 'February'),vals)

df$vals <- round(df$vals, 3)



ggplot(df, aes(x=Year, y=vals))+geom_bar(stat='identity', position='dodge') + geom_text(aes(label=vals), vjust=-0.25)+
  xlab("Year")+ylab("Predictive Strength")+theme_minimal()+ggtitle("Logit Model Accuracy")


```





```{r message=FALSE, warning=FALSE, include=FALSE}

## Odds Ratios
exp(coef(model_2019_blocks))
exp(coef(model_2021))
exp(coef(feb_model))

exp(cbind(OR = coef(model_2019_blocks), confint(model_2019_blocks)))
exp(cbind(OR = coef(model_2021), confint(model_2021)))
exp(cbind(OR = coef(feb_model), confint(feb_model)))


```




## Odds Ratio Box Plot


```{r echo=FALSE, message=FALSE, warning=FALSE}
boxLabels = c("Hispanic", "White", "Black", "Asian")

##2019


df <- data.frame(lables=boxLabels, 
                  boxOdds = c(1.0021304779, 0.9934778163  , 1.0200081806 ,1.0312674226    ), 
                        boxCILow = c(1.0014661525   , 0.9931773600   ,1.0195372923 ,1.0295998976     ), 
                        boxCIHigh = c(1.002793072, 0.993778233, 1.020480101,1.032950946))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("2019 Model Odds Ratios, Race")


##2021


df <- data.frame(lables=boxLabels, boxOdds = c(1.012368848609690   , 
0.993498405513978   , 1.013385901077802,1.042262592003207    ), boxCILow = c(1.011481634734796   , 0.993044840129250   ,1.012728356907708,1.040464698203379   ), 
                        boxCIHigh = c(1.013251325461511, 0.993952139817316, 1.014042344750129,1.044072940422991))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("2021 Model Odds Ratios, Race")



##February


df <- data.frame(lables=boxLabels, boxOdds = c(1.00953308    , 
0.99998742    , 1.00236907 ,1.02191628     ), boxCILow = c(1.0088145    , 0.9996327    ,1.0018087 ,1.0203451    ), 
                        boxCIHigh = c(1.01024899, 1.00034257, 1.00292776,1.02348719))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("February Model Odds Ratios, Race")


# Median Income


boxLabels = c("2019 District", "2021 Districts", "February")

df <- data.frame(lables=boxLabels, boxOdds = c(2.0259539656   , 5.310856286752776 , 1.18227618    ), boxCILow = c(1.9723423285   , 5.108356308166039 ,1.1563041   ), 
                        boxCIHigh = c(2.081119204,5.521796860873518,1.20975856))

p <- ggplot(df, aes(x = boxOdds, y = boxLabels)) 


ggplot(data = df,
       mapping = aes(y = forcats::fct_inorder(f = (x = boxLabels)))) +
  geom_vline(xintercept = 1) +
  geom_point(mapping = aes(x = boxOdds)) +
  geom_errorbarh(mapping = aes(xmin = boxCILow,
                               xmax = boxCIHigh)) +
  coord_trans(x = scales::exp_trans()) +
  # scale_x_continuous(breaks = log(x = 0.5 * (1:10)),
  #                    minor_breaks = NULL,
  #                    labels = (0.5 * (1:10))) +
  labs(x = "Odds Ratio",
       y = "") +theme_minimal() + ggtitle("Median Income Logged")




```


## Random Forest 


```{r echo=FALSE, message=FALSE, warning=FALSE}


model_data <- centroids %>% 
  select(PARTY_2019,PARTY_2021,PARTY_2022, PCT_BLC, PCT_HSP, PCT_WHT, PCT_ASN, Med_income_log)
model_data$PARTY_2021 <- as.factor(model_data$PARTY_2021)
model_data$PARTY_2019 <- as.factor(model_data$PARTY_2019)
model_data$PARTY_2022 <- as.factor(model_data$PARTY_2022)

samp <- sample(nrow(model_data), 0.8*nrow(model_data))

train <- model_data[samp,]
test <- model_data[-samp,]

RF_2019 <- randomForest(
  formula = PARTY_2019 ~ PCT_BLC + PCT_HSP+PCT_WHT+PCT_ASN + Med_income_log,
  data=train,
  importance=TRUE
)

# importance(RF_2019)
varImpPlot(RF_2019)
#=====================================================================#


RF_2021<- randomForest(
  formula = PARTY_2021 ~ PCT_BLC + PCT_HSP+PCT_WHT+PCT_ASN + Med_income_log,
  data=train,
  importance=TRUE
)

# importance(RF_2021)
varImpPlot(RF_2021)
#=====================================================================#


RF_Feb<- randomForest(
  formula = PARTY_2022 ~ PCT_BLC + PCT_HSP+PCT_WHT+PCT_ASN + Med_income_log,
  data=train,
  importance=TRUE
)

# importance(RF_2021)
varImpPlot(RF_Feb)









```




Finally, lets see the predictive fit for all three. 



```{r echo=FALSE, message=FALSE, warning=FALSE}



p <- predict(RF_2019, newdata=test)
# table(p, test$PARTY_2019)

results <- cbind(p, test$PARTY_2019)

colnames(results) <- c('pred','real')
results <- as.data.frame(results)

pred_2019 <- sum(p==test$PARTY_2019) / nrow(test)


#=====================================================================#



p <- predict(RF_2021, newdata=test)
# table(p, test$PARTY_2021)

results <- cbind(p, test$PARTY_2021)

colnames(results) <- c('pred','real')
results <- as.data.frame(results)

pred_2021 <- sum(p==test$PARTY_2021) / nrow(test)


#=====================================================================#


p <- predict(RF_Feb, newdata=test)
# table(p, test$PARTY_2021)

results <- cbind(p, test$PARTY_2022)

colnames(results) <- c('pred','real')
results <- as.data.frame(results)

pred_feb <- sum(p==test$PARTY_2022) / nrow(test)


#=====================================================================#



vals <- c(pred_2019, pred_2021,pred_feb)

df <- data.frame(Year = c('2019', '2021', 'February'),vals)

df$vals <- round(df$vals, 3)



ggplot(df, aes(x=Year, y=vals))+geom_bar(stat='identity', position='dodge') + geom_text(aes(label=vals), vjust=-0.25)+
  xlab("Year")+ylab("Predictive Strength")+theme_minimal()+ggtitle("Random Forest Model Accuracy")

```




## Conclusion




With the metrics developed here, the February congressional districts seem to be more fair compared to the lines struck down by the NC Supreme Court. Relativity is important to recall, as the high degree of likelihood that political lines are drawn around demographic lines are clearly demonstrated. The February update is an improvement on the abomination of the lines originally approved by the NCGA, but that should not be mistaken for progress. 


















































