---
title: "Quantifying NIMBYism"
author:
  - name: Jacob Ford
    url: https://jford-dchcmpo.github.io/Blog/
date: 12-08-2021
output:
  distill::distill_article:
    self_contained: false
---

## Vapid & Plastic

Take a stroll down any street in Durham, which many on this blog have called the Mecca of the South, and an ubiquitous lawn ornament is the variety of signs. Black lives matter. Science is Real. Housing is a Human Right. Stop the Steal. 

No, I'm kidding about the last one, at least thoe aren't prevelant in my neighborhoods. 

Those signs are virtue signaling in my opinion. But I'm not here to argue about lawn signs. I found myself wondering, are Durham zoning laws reflective of a majority Democratic enclave, a quickly growing liberal bubble? 


Are Durham's zoning ordinances aligned with beliefs of all those stupid lawn signs? 



```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(caliperR)
library(nngeo)
library(tcadr)
library(sf)
library(leaflet)
library(PL94171)
library(janitor)
library(dplyr)
library(tidycensus)
library(tidyverse)
options(tigris_use_cache = TRUE)
library(sf)
library(data.table)
library(dplyr)
library(plotly)


# remotes::install_github("walkerke/tidycensus")
# remotes::install_github("walkerke/tigris")
library(tidycensus)
library(tidyverse)
library(tigris)

vars_2020 <- load_variables(2020,"pl",cache=TRUE)
vars_2010 <- load_variables(2010,"sf1", cache=TRUE)
vars_2019 <- load_variables(2019,"acs5", cache=TRUE)




zones <- st_read("C:/Users/JacobFo/OneDrive - City of Durham/GitHub/Personal Blog/Personal Blog/_posts/2021-12-08-Quantifying-NIMBYism/Zoning.shp")



summary(zones)




```
 
 
 
## EDA 

Durham's zoning data is publicly available, and is classified in an understandibly complex manner. Below are the initial zone types: 

```{r echo=FALSE, message=FALSE, warning=FALSE}
table(zones$ZONE_GEN)
```
As you can see, a lot of these can be collapsed. 


```{r}
res_words <- c("RESIDENTIAL", "RES")

commerical_words <- c("COMMECIAL", "COMMERCIAL", "UNIVERSITY","OFFICE", "INDUST", "MIXED", 'COMPACT')
  



zones <- zones %>%
  mutate(
    com_flag = case_when(
      grepl(commerical_words, ZONE_GEN) ~ 1,
      ),
    res_flag=case_when(
      grepl(res_words, ZONE_GEN) ~ 1,
        )
    )
```


```{r echo=FALSE, message=FALSE, warning=FALSE}

pal <- colorFactor(
  palette='Dark2',
  domain=zones$ZONE_GEN
)

zone_map <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-78.90390102117877,35.998220055791876,  zoom=14)%>%
  addPolygons(data=zones,
              weight=1,
              color=~pal(ZONE_GEN),
              popup=paste("Zone Type: ", zones$ZONE_GEN)) %>%
  addLegend(
    position="bottomright",
    pal=pal,
    values=zones$ZONE_GEN
  )


zone_map
```







## References

1. https://medium.com/@NYUurbanlab/actually-nimbys-cities-arent-building-enough-51806be038ae
2. https://abundanthousingla.org/meet-l-a-countys-biggest-nimbys/
3. https://live-durhamnc.opendata.arcgis.com/datasets/zoning/explore?location=36.012451%2C-78.920767%2C15.62

4. https://maps.durhamnc.gov/?x=36.01483037908998&y=-78.92072715568013&z=18055.9548215&r=0&b=11&pid=NA&s=zoning&l=active_address_points,countymask,transitionalofficeoverlay,NPO,NHD,LocHistLandmarks,airportoverlay,citylimits,RTPboundary,parcels,zoning,



